---
title: "car_viz"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tinytex)
#have to call tinytex in order to knit to pdf
```
# call built-in data mtcars.
```{r}
data(mtcars)
```
# Select only car models where mpg<20
```{r}
mtcars_mpg2 <- mtcars[mtcars$mpg < 20,]
#use of $ to select the desired column
```

# Reduce the variables to mpg, cyl, disp, hp, gears
```{r}
mtcars_mpg2 <- mtcars_mpg2[, c(1,2,3,4,10)]
#columns are vectors so we are calling the columns we want by the number vector they are in the data frame
```
# read the R file hand_functions.R so that it can be used
# notice that with echo = TRUE 
````{r}
source(file = "hand_functions.R", echo = TRUE)
```
# Now use the function from hand_functions.R
```{r}
sp_out <- sum_special(mtcars_mpg2)

sum_special <- function(df_x){
  
  ## sum_special calculates data summary statistics
  ## the input param df_x is the data frame of input values
  
  #  browser()  # browser() will start the debugger
               # if the line is uncommented
  
  
  ## test the input data to assure that it is a data frame.
  try(if(!is.data.frame(df_x)) stop("Input data must be a data frame."))
  
  
  sp_means <- apply(df_x, MARGIN = 2, FUN = mean)
  sp_var <- apply(df_x, MARGIN = 2, FUN = var)
  sp_cov <- cov(df_x)
  sp_cor <- cor(df_x)
  
  ## Note that defining a list with the 
  ## syntax list(list_name = list_content) produces
  ## named list items
  sp_outputs <- list(sp_means=sp_means, 
                     sp_var = sp_var, 
                     sp_cov = sp_cov, 
                     sp_cor = sp_cor)
  
  return(sp_outputs)
}

```

```{r}
#call the sp_out list

sp_out

#This shows you the mean, variance, covariance, and correlation between the variables mpg, cyl, disp, hp, and gear.
```


```{r}
# library(esquisse)
# 
# esquisser(data = mtcars_mpg2, viewer = "browser")

#Using esquisser allows us to easily manipulate the plot without having to worry about the code. Once we have the desired plot we can copy the ggplot code into rstudio. Esquisser can be used in a own browser window, which makes looking back and forth between r and the esquisser window very easy.
```

```{r}
ggplot(mtcars_mpg2) +
  aes(x = disp, y = mpg) +
  geom_point(shape = "bullet", size = 4L, colour = "#B22222") +
  geom_smooth(span = 0.5) +
  theme_minimal()
```

# note that this boxplot cannot be made with esquisse() unless
# the data is adjusted.  What adjustment is needed?
# We need to use cylinders as the x input rather than disp, and used the data set mtcars_mpg2 which is a subset of mtcars that only includes car models where mpg<20.

```{r}
ggplot(mtcars_mpg2, aes(x=as.factor(cyl), y=mpg)) + 
  geom_boxplot(fill="slateblue", alpha=0.2) + 
  xlab("cyl")
```





